### 2024-11-21 paper.tex
p inf proof draft

One of the emergent features in UniswapV3 is that you can use concentrated
liquidity positions as limit orders, with the problem that, due to their
logarithmic nature, as prices grow, tick precision decreases. To give
an example, if the current price of bitcoin was 80000USDC, the difference
between the 2 closest ticks is of almost 8USDC. For this reaon, Osmosis introduces
geometric tick spacing with additive ranges \cite{osmosis-price-function}, that
allows us to define an order book on top of the AMM with proper tick precision.
Its easy to verify that its price function is:
\[
\rho(t) = 10^{\lfloor t/9e6\rfloor - 6} \left(
	t+10^6\left(1 - 9e6\left\lfloor \frac t{9e6}\right\rfloor
	\right)
\right)
\]

@@

\section{Price function inverse}
The only missing piece is the generalized inverse $\rho^{-1}$ that will let
us map every price to its closest tick in Osmosis. We will prove that such
function is:
\[
\rho^{-1}(p) = 10^{6-\lfloor\log p\rfloor} \left(
	p+10^{\lfloor\log p\rfloor}\left(
		9\lfloor\log p\rfloor-1
	\right)
\right)
\]
Of course, it could be the case that there was 2 different prices with
the same tick, under the inverse. Dualy, there could be 2 ticks that map
to the same price. Thats why by generalized inverse we mean 2 things:
\begin{enumerate}
    \item Generalized inverse property: $\forall t\in T: \rho^{-1}(\rho(t)) =
        t$, where $T\subset\mathbb Z$ is the tick space.
    \item Closeness property: $\forall t_0, t_1\in T$ st $t_1 > t_0$, we have
        that $\forall p\in [\rho(t_0), \rho(t_1)]$ its the case that
        $\rho^{-1}(p)\in[t_0, t_1]$. That is, given any 2 ticks, if we map them
        to the price space and pick any price in that range, that price tick will
        also be in the tick range after pulling it back.
\end{enumerate}
But its easy to check that the weaker form of the closeness property when $t_1
= t_0 + 1$ is equivalent to the general form. And by dualizing the weaker form,
that is, taking ranges of the form $[t_1 - 1, t_1]$, we can trivially prove that
the generalized inverse property holds. Thus, we only need to prove the weaker
closeness property to prove that $\rho^{-1}$ is properly defined.

%But still work in progress; The derivation is not important, whats important
%is to prove its a generalized inverse of osmosis price function.
%\subsection{Computation security proof}
%Work in progress. I already got the proof in a obsidian note, just need
%to review and formalize it.

>>>>

One of the emergent features in UniswapV3 is that you can use concentrated
liquidity positions as limit orders, with the problem that, due to their
logarithmic nature, as prices grow, tick precision decreases. To give
an example, if the current price of bitcoin was 80000USDC, the difference
between the 2 closest ticks is of almost 8USDC. For this reaon, Osmosis introduces
geometric tick spacing with additive ranges \cite{osmosis-price-function}, that
allows us to define an order book on top of the AMM with proper tick precision.
Its easy to verify that its price function is:
\[
\rho(t) = 10^{\lfloor t/9e6\rfloor - 6} \left(
    t - 10^6\left(9\lfloor\frac{t}{9e6}\rfloor - 1\right)
\right)
\]

@@

\section{Price function inverse}
The only missing piece is the generalized inverse $\rho^{-1}$ that will let
us map every price to its closest tick in Osmosis. We will prove that such
function is:
\[
\rho^{-1}(p) = 10^{6-\lfloor\log p\rfloor} \left(
	p+10^{\lfloor\log p\rfloor}\left(
		9\lfloor\log p\rfloor-1
	\right)
\right)
\]
Of course, it could be the case that there was 2 different prices with the same
tick, as we require ticks to be integers (ie, we will take $\rho^{-1}(p) :=
\lfloor \rho^{-1}(p)\rfloor$). Moreover, there could be 2 ticks that map to the
same price, due to computational precision. Thats why by generalized inverse we
mean 2 things:
\begin{enumerate}
    \item Generalized inverse property: $\forall t\in T: \rho^{-1}(\rho(t)) =
        t$, where $T\subset\mathbb Z$ is the tick space.
    \item Closeness property: $\forall t_0, t_1\in T$ st $t_1 > t_0$, we have
        that $\forall p\in [\rho(t_0), \rho(t_1)]$ its the case that
        $\rho^{-1}(p)\in[t_0, t_1]$. That is, given any 2 ticks, if we map them
        to the price space and pick any price in that range, that price tick will
        also be in the tick range after pulling it back.
\end{enumerate}
But its easy to check that the weaker form of the closeness property when $t_1
= t_0 + 1$ is equivalent to the general form. The generalized inverse property
is not enough for our usecase, as we just mentioned there will be prices that
map to the same tick. Informally, by proving the weak generalized closeness
property, we will know that any price will get mapped to its closest tick or
its second closes tick (which, for our usecase, is good enough). The proof
sketch goes as follows:
\begin{enumerate}
    \item Prove the generalized inverse property by simple brute force (as the
        domain is finite and relativelly small).
    \item Observe both $\rho^{-1}$ and $\rho$ are monotone non-decreasing.
    \item Conclued trivially that the weak closeness property holds, by simply
        taking inverses over $\rho(t_0) \leq p \leq \rho(t_0 + 1)$.
\end{enumerate}

\subsection{Computation security proof}
Once again, we will have to prove that $\rho^{-1}$ computation is secure.
In the codebase:

\begin{minted}[fontsize=\small]{rust}
let compute_price_inverse = |p: &Decimal| -> Option<i32> {
    let floor_log_p = PositiveDecimal::new(p)?.floorlog10(); // (1)
    let x = floor_log_p.checked_mul(9)?.checked_sub(1)?;     // (2)

    let x = maybe_neg_pow(floor_log_p)?                      // (3)
        .checked_mul(SignedDecimal256::new(x.into())).ok()?  // (4)
        .checked_add(p.clone().try_into().ok()?).ok()?;      // (5)

    let x = maybe_neg_pow(6i32.checked_sub(floor_log_p)?)?   // (6)
	    .checked_mul(x).ok()?;                           // (7)

    let x: Int128 = x.to_int_floor().try_into().ok()?;       // (8)
    x.i128().try_into().ok()                                 // (9)
};
\end{minted}
Thus, we want to prove the closure will never return \texttt{None} at
each step:
\begin{enumerate}[label=(\arabic*).]
    \item \texttt{floorlog10} computes $\lfloor\log p\rfloor$. The implementation is:
\begin{minted}[fontsize=\small]{rust}
pub fn floorlog10(self: &PositiveDecimal) -> i32 {
    let x: u128 = self.0.atomics().into();
    // Invariant: `u128::ilog10(u128::MAX)` fits in `i32`.
    let x: i32 = x.ilog10().try_into().unwrap();
    // Invariant: `ilog10(1) - 18 = 0 - 18` fits in `i32`.
    x.checked_sub(18).unwrap()
}
\end{minted}
And as we can see, its security is already proved in the comments.
Note that we subtract $18$ to our result, as we computer the decimal
logarithm of the atoms of a \texttt{Decimal} type (wrapped trivially
in \texttt{PositiveDecimal}), thus we have to compensate for the $18$
decimal points it uses. Its also trivial to verify that 
\[
\text{Im}(\texttt{floorlog10}) =[-18, 20]\subset\mathbb Z
\]
is invariant.
    \item Taking the upper and lower bound of the image of \texttt{floorlg10},
        we get the valid \texttt{i32}:
        \[
            (9\lfloor\log p\rfloor - 1) \in[-162, 179]
        \]
    \item \texttt{maybe\_neg\_pow} sipmply takes \texttt{SignedDecimal256::pow(10, exp)}
        but with \texttt{i32} domain instead of \texttt{u32} (the implmentation is trivial).
        Clearly, the operation could fail for the whole domain, but in our case:
        \begin{equation}
        \begin{split}
        &\texttt{maybe\_neg\_pow}(\text{Im}(\texttt{floorlog10}))=\\
        &\texttt{maybe\_neg\_pow}([-18, 20])
        = [10^{-18}, 10^{20}] \subset\texttt{SignedDecimal256}
        \end{split}
        \end{equation}
        Note that were using the \texttt{Signed} version of \texttt{Decimal256}
        to compute $10^{\lfloor\log p\rfloor}$ as were about to multiply its
        result by the most inner term of $\rho^{-1}$ (that is, $9\lfloor\log
        p\rfloor - 1$, which we already know could be negative).
    \item We got $(9\lfloor\log p\rfloor - 1) \in[-162, 179]$ and
        $10^{\lfloor\log p\rfloor}\in[10^{-18}, 10^{20}]$. Thus,
        the product could only overflow if $179\cdot 10^{20}$ could not fit
        in \texttt{SignedDecimal256}, but thats clearly not the case.
    \item We compute $p + x$, where $x$ is the term computed in step $(4)$,
        that is, $179\cdot 10^{20}$ in the worst case. Thus, this addition
        could only overflow if $\texttt{u128::MAX} + 179\cdot 10^{20} = 2^{128}
        - 1 + 179\cdot 10^{20}$ did not fit in \texttt{SignedDecimal256}, but
        thats clearly not the case.
    \item WIP
    \item WIP
    \item WIP
    \item WIP
    %\item Computamos el término $10^{6-\lfloor\log p\rfloor}$. Tenemos: $$ (6 -
    %    \lfloor\log p\rfloor) \in [6 - 20, 6 -(-18)] =[-14, 24] $$ Luego
    %    cláramente $10^{24}$ cabe en \texttt{SignedDecimal256}. Igualmente, tanto
    %    $-14$ como $24$ caben en \texttt{i32} de sobra.
\end{enumerate}
